---
title: "Heart Disease"
author: "Munther Alnghemshi"
date: "11/19/2020"
html_document:
  toc: yes
  toc_depth: '3'
  df_print: paged
---
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Packages
library(tidyverse)
library(ggplot2)
library(readr)
library(ROCR)
library(caret)
library(gbm)
library(corrplot)
library(MASS)
library(rpart)
library(caTools)
library(class)
library(glmnet)
library(Hmisc)
library(pROC)
library(scales)
library(cluster)
library(factoextra)
library(corrplot)
library(ggpubr)
library(DT)
 

```
 

\
\
\
\
\
\
\
\
\
\


# Introduction:

_This data contain information related to factor responsible for Heart Attack, so our goal is to analyze the trends to predict a pattern or indications that could lead to a heart attack._

## Attribute Information:
_This database contains 13 attributes_
 

| Variable | Description |
|----------|-------------|
|`Age`||
|`Sex`| Male = 1, Female = 0 |
| `Chest Pain` | 4 Values: Value 1: typical angina Value 2: atypical angina Value 3: non-anginal pain Value 4: asymptomatic  |
| `trestbps` | Resting blood pressure (in mm Hg on admission to the hospital) |
| `chol` | Serum cholestoral in mg|
|`fbs`|(fasting blood sugar > 120 mg/dl) (1 = true; 0 = false)|
|`restecg`|resting electrocardiographic results (values 0,1,2)|
|`thalach`|Maximum heart rate achieved|
|`exang`|Exercise induced angina (1 = yes; 0 = no)|
|`oldpeak`|ST depression induced by exercise relative to rest|
|`slope`|The slope of the peak exercise ST segment ( Value 1: upsloping, Value 2: flat, and Value 3: downsloping)|
|`ca`|Number of major vessels (0-3) colored by flourosopy|
| `thal`   | 3 = normal; 6 = fixed defect; 7 = reversable defect |
 \
 \
 \
 \
 \
 \
 \
 
## Data Overview:
 _Below we can see the data table showing all attributes_
```{r, echo=FALSE, warning=FALSE, message=FALSE}
data <- read_csv("HeartD.csv")
datatable(data)#showing the data
 
 
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
 
```

***
### Averages:
```{r echo=FALSE, warning=FALSE, message=FALSE}
cws <- data %>%
  group_by(sex) %>%  #taking the avg by sex
  summarise(NumberofPatients  = n(),
            Avg_Age=mean(age), 
            SD_Age= sd(age), 
            Avg_ChestPain= mean(cp), 
            Avg_trestbps= mean(trestbps), 
            Avg_chol= mean(chol), 
            Avg_thalach=mean(thalach))
datatable(cws, caption = "All averages for the given values")  


```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
ra=range(data$age)
me=mean(data$age)
sd=sd(data$age)
va=var(data$age)
co=cor(data$age,data$target)

```
```{r, echo=FALSE, warning=FALSE, message=FALSE}

#interpreting the values with respect to the first table above
#cpoied the below code from kaggle: https://www.kaggle.com/utkarsh1997/heart-disease-prediction
D2 <- data %>% 
  mutate(sex = if_else(sex == 1, "MALE", "FEMALE"), #if sex is 1 then conver to male 
         fbs = if_else(fbs == 1, ">120", "<=120"),
         exang = if_else(exang == 1, "YES" ,"NO"),
         cp = if_else(cp == 1, "ATYPICAL ANGINA",
                      if_else(cp == 2, "NON-ANGINAL PAIN", "ASYMPTOMATIC")),
         restecg = if_else(restecg == 0, "NORMAL",
                           if_else(restecg == 1, "ABNORMALITY", "PROBABLE OR DEFINITE")),
         slope = as.factor(slope),
          
         ca = as.factor(ca),
         thal = as.factor(thal),
         target = if_else(target == 1, "YES", "NO")
         ) %>% 
  mutate_if(is.character, as.factor) %>% 
  dplyr::select(target, sex, fbs, exang, cp, restecg, slope, ca, thal, everything())
```
***

## Age Analysis:

1. The age range goes form 34 up to 76 years old, and the average age is 55.7\



```{r, echo=FALSE, warning=FALSE, message=FALSE}
boxplot(data$age, col = "yellow", horizontal = TRUE, main= "Box Plot of Age") #showing the avg age of population

```



2. The SD is 9.41, which means that $95%$ of the population lay between the age group of 37 and 75 years\
```{r, echo=FALSE, warning=FALSE, message=FALSE}
hist(data$age, main = "Distribution of age", xlab= "Age", ylab = "Counts/Frequency", labels = TRUE)
 
```


3. There is a negative correlation $-0.207$ between age and target meaning that the higher the age number is the lower the probability of having a heart attack with respect to the given data set. 
```{r, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(data,aes(x=age,y=target)) + geom_smooth()+ geom_point()+  ggtitle("Curve showing relation between Age and Target")
 
```

*** 

## Sex Analysis:

1.Approx 68% of the population is male and 32% is female \
```{r, echo=FALSE, warning=FALSE, message=FALSE}
percent=round(table(data$sex)/sum(table(data$sex))*100)
 
paste=paste(percent,"%",sep=" ", c("Female","Male"))
pie <- pie(table(data$sex), labels=paste, main = "Percentage of Male to Female")

```
\
\
2. 207 out of 303 patients are males and the rest are females\

```{r, echo=FALSE, warning=FALSE, message=FALSE}
bar <- barplot(table(D2$sex),
        xlab="Gender",
        ylab="Frequency",
        main="Bar Plot of Sex")
```



3. We can conclude from the visualized data below that male are more likely to have a heart attack campared to felame population\


 
 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
 
ggplot(D2,aes(x=sex,fill=target))+
  geom_bar(alpha=.5)+
  ggtitle("Sex count vs Heart attack presence")
```
***

## Chest pain:
1. 0-pain level is the most recorded value, after that comes pain-2 level
 
```{r}
ggplot(data, aes(x=cp,fill=target))+
  geom_bar() 
  
```
2. The average chest pain level is almost 1. 
```{r}
boxplot(data$cp, col = "red", horizontal = TRUE, main= "Box Plot of chest pain")

```

3. Patient with pain-2 level are more likely to have a heart attack 


 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#finding wich pain level is likely to cause a H.A
target1 <- data %>%  
  filter(data$target==1) 
ggplot(target1, aes(x=cp,fill=target))+
  geom_bar()+
  ggtitle("Pain level in the case of a heart attack")


```
4. Taking the fact that patiants with pain-2 level are more likely, the histo below shows the distribution of age with chest pain level 2 concluding that people from the age 50-55 are more likely to get a heart attack in the case of a chest pain level 2 (This pain level is more likely to cause a heart attack as explained above)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
dd <- data %>% 
  filter(data$cp==2)
hist(dd$age, main = "Distribution of age with level pain =2", xlab= "Age", ylab = "Counts/Frequency", labels = TRUE)
```

***

## Maximum heart rate:

1.We can see below that as maximum heart rate increases a patient will be brone to a heart attack. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
 
plot=ggplot(data,aes(x=thalach,y=target))+geom_point()


plot+geom_smooth()+
    scale_x_continuous(name="thalach")+
    scale_y_continuous(name="Probability of Heart Attack")+
    ggtitle("Relation Between Heart rate and getting a Heart attack")
```


## 	Exercise induced angina:
As suggested by the curve below, as exang increases --> the chance of a heart attack increase. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
plot=ggplot(data,aes(x=exang,y=target))+geom_point()


plot+geom_smooth()+
    scale_x_continuous(name="exang")+
    scale_y_continuous(name="Probability of Heart Attack")+
    ggtitle("Relation Between Exercise induced angina and getting a Heart attack")
```


## OldPeak:
